---
title: "Jenn workspace"
output: html_document
---
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
library("stringr")
library(readr)
library(data.table)
library(ggmap)

```

```{r}

#bigfoot_base <- read.csv("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/bis15W2025_Group11-main/data/bfro_updatedData2.csv")

#bigfoot <- bigfoot_base %>%
 # select(county, state,season,  title,  latitude,  longitude, date,  number,  classification,  temperature_high,   temperature_mid, temperature_low,  dew_point,  humidity,  cloud_cover,  moon_phase,  precip_intensity,  precip_probability,  precip_type,   pressure,   summary,  conditions,  uv_index,  visibility,   wind_bearing, wind_speed) %>%
#  filter(state=="California") %>%
#  filter(latitude != "NA")%>%
#  filter(longitude != "NA")
  
#write.csv(bigfoot, "C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/bigfoot_updated_CAonly_noNA.csv")

bigfoot <- read.csv("data/bigfoot_updated_CAonly_noNA.csv")
```


#sightings map
```{r}
bigfoot <- read.csv("data/bigfoot_updated_CAonly_noNA.csv")



bigfoot %>% 
  select(latitude, longitude,) %>%  #f command is the buffer. 
  summary()

```

```{r}
register_stadiamaps("35c776b4-832b-4b73-bbec-f195a0bdbb2c", write = FALSE)
lat <- c(32.72, 41.99)
long <- c(-124.2, -116.5)
bbox <- make_bbox(long, lat, f = 0.03)

bigfootmap <- get_stadiamap(bbox, maptype = "stamen_terrain", zoom=7) #zoom impacts how large the map is in memory


ggmap(bigfootmap)
```

```{r}
map<- ggmap(bigfootmap) + 
  geom_point(data = bigfoot, aes(longitude, latitude), size=0.4) +
  labs(x= "Longitude", y= "Latitude", title="Bigfoot sightings") +
  ggsave("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/sightings.png")


```

```{r}
#weather work

bigfoot %>% 
  count(summary) %>%
  arrange(-n)

bigfoot %>% 
  filter(summary != "")%>%
count(summary) %>%
  arrange(-n)%>%
  pivot_wider(names_from = "summary", values_from = "n") %>%
 summarise(across(contains("cloudy"), ~ sum(.x, na.rm = TRUE)))


```

```{r}
#sightings by county, calc percentage of all CA sightings
bigfoot %>%
  count(county)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_CA_sightings = (n/total)*100)        
          

bigfoot %>%
  ggplot(aes(x=county))+
  geom_bar()+
  coord_flip()+
  labs(title = "Bigfoot Sightings by County", x="County", y="Count")


```



```{r}
#sighting classifications



bigfoot %>%
  ggplot(aes(x=classification, fill=classification)) + 
  geom_bar()  +     #good for counts
  labs(title = "Bigfoot Sightings by Classification", x="Sighting classification", y="Count") #+

   #ggsave("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/classifications.png", width=4, height=4)

```

```{r}
#bigfoot by month
 
bigfootDated <- bigfoot%>%
  separate(date, into= c("month", "day", "year"), sep = "/")

#rename variables in month column to month name


bigfootDated %>%
  count(year,classification)%>%
  top_n(20, n) %>% 
  ggplot(aes(x=year, y=n, fill=classification))+
  geom_col() +  #specifies an x and y 
    coord_flip()+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))+
  labs(title = "Years with most bigfoot sightings", x="count", y="year")
# ggsave("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/top20years.png", width=4, height=4)


bigfootDated %>%
  count(month)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_month = (n/total)*100) 

bigfootDated %>%
  count(day)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_day = (n/total)*100) 

```

```{r}
bigfootDated %>% 
  filter(conditions != "")%>%
count(conditions) %>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_weather = (n/total)*100) 



```



```{r}
#sightings by date

#glimpse(bigfoot)

bigfootDated %>%
  count(year,classification)%>%
  top_n(20, n) %>% 
  ggplot(aes(x=year, y=n, fill=classification))+
  geom_col() +  #specifies an x and y 
    coord_flip()+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))+
  labs(title = "Years with most bigfoot sightings", x="count", y="year")
# ggsave("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/top20years.png", width=4, height=4)




bigfootDated %>%
  filter(year != "NA")%>%
  count(month, day,classification)%>%
  ggplot(aes(x=month, y=n, fill=classification))+
  scale_x_discrete(limits= c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  geom_col() +
  labs(title = "months with most bigfoot sightings", x="month", y="Number of sightings")
 # ggsave("C:/Users/Jennifer/Documents/GitHub/bis15W2025_Group11/data/months.png", width=6, height=4 )


```

```{r}


bigfootDated %>%
 count(county)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_CA_sightings = (n/total)*100)    

SBdays <- bigfootDated %>%
  count(day)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_day = (n/total)*100)

bigfootDated %>%
  count(month)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_month = (n/total)*100)

  
SBconditions <- bigfootDated %>%
  filter(conditions != "")%>%
  count(conditions) %>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  mutate(perc_sightings_weather = (n/total)*100) 


SB <- tibble(SBcounties,SBdays)

```

```{r}
alcohol_CA <- read.csv("alcohol_consumption.csv")%>%
  clean_names()%>%
  filter(state=="California")

alcohol_CA$spacer <- "County"

glimpse(alcohol_CA)

alcohol_CA_clean<- alcohol_CA %>%
    unite(county, "county", "spacer", sep=" ")
    
alcohol_CA_clean%>%
  rename(alcohol_CA_clean, "excessive_alc"=x..excessive.drinking)

glimpse(alcohol_CA_clean)
glimpse(bigfootDated)

 
#right_join(bigfootDated, alcohol_CA_clean, by="County")
bigfoot_alc <- full_join(bigfootDated, alcohol_CA_clean, by="county")

bigfoot_alc %>% 
  count(county)%>%
  ggplot(aes(y=n, x=Excessive_alc))+ #notice the switch in x and y
  geom_col()

```


## Adding Numeric Input
Let's try adding a new numeric input so the user can change the size of the points on the plot with a slider. We can do this with the `sliderInput()` function. The `sliderInput()` function takes a label that the server function will use to access the value, the label that the ui will display to the user, a minimum and maximum value, a starting value, and a step size. Edit the script to include the slider input in the ui and for the server to access the input value.
```{r, eval=F, echo=T}
testrun<- read.csv("data/testrun.csv")

      v<- testrun%>%
          summarize(mean_val=mean(value))


    testrun %>% 
      ggplot(aes(x=param, y=value)) + 
      geom_col() +
         labs(x = "parameter", y = "percent of total sightings in california by parameter")+
     ggtitle( paste("Calculated chance of seeing bigfoot ", v, "%"))

    
    
    
param<- c(1,2,4)
value<- c(12,3,5)
trial<- tibble(param, value) 

trial%>%
      ggplot(aes(x=param, y=value)) + 
      geom_col() +
         labs(x = "parameter", y = "percent of total sightings in california by parameter")+
     ggtitle( paste("Calculated chance of seeing bigfoot ", v, "%"))    
    


```


```{r}
#library(shiny) 
ui <- dashboardPage(
 

   
  dashboardHeader(title = "SasqWATCH"),
  
  dashboardSidebar(disable=T),
  
  dashboardBody(
  
    fluidRow(
      box(title = "Plot Options", 
          width = 3, #adjust the column width
              
                #input I'm trying to pass the input variables into a vector/df to use for the graph later
                selectInput("a", 
                            "Select County", 
                            choices = unique(bigfootDated$county), #instead of typing all options, we can use unique()
                            selected="Tuolumne County"),
                selectInput("b", 
                            "Select Month", 
                            choices = c(1:12),
                            selected="1"),
                
                selectInput("c", 
                            "Select day", 
                            choices = c(1:31), 
                            selected="1"),
                plotOutput(

                  "plot", width="500px", height="400px"),
verbatimTextOutput(input$a)
        
), # close the first box    




box(title = "Plot of Penguins Data", width = 6, #adjust the column width and add title
  plotOutput("plot", width = "500px", height = "400px")
) # close the second box
) # close the row
) # close the dashboard body
) # close the ui

    
server <- function(input, output, session) {
  
  session$onSessionEnded(stopApp) # stop the app when we close it
  
  output$plot <- renderPlot({
  user_vector <- reactiveValues()
  user_param<- input$a

value<-c(a,5,7)

value$a <- bigfootDated %>%
  count(county)%>%  
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter( county == "Amador County")%>%
  mutate(perc_CA_sightings = (n/total)*100)    

value$b <- bigfootDated %>%
  count(day)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter(day== "7")%>%
  mutate(perc_sightings_day = (n/total)*100)

value$c <- bigfootDated %>%
  count(month)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter(month=="8")%>%
  mutate(perc_sightings_month = (n/total)*100)


param<- c(input$a,input$b,input$c)
value<- c(value$a,value$b,value$c)

print(param)
print(value)


trial<- tibble(param, value) 

#trial %>% 
  #mutate(across(param, value, as.double))

      v<- trial%>%
          summarize(mean_val=mean(value))

trial%>%
      ggplot(aes(x=param, y=value)) + 
      geom_col() +
      labs(x = "parameter", y = "percent of total sightings in california by parameter")+
     ggtitle( paste("Calculated chance of seeing bigfoot ", "XX", "%"))    

    })
}

shinyApp(ui, server)
```



```{r}
#library(shiny) 
ui <- dashboardPage(
 

   
  dashboardHeader(title = "SasqWATCH"),
  
  dashboardSidebar(disable=T),
  
  dashboardBody(
  
    fluidRow(
      box(title = "Plot Options", 
          width = 3, #adjust the column width
              
                #input I'm trying to pass the input variables into a vector/df to use for the graph later
                selectInput("a", 
                            "Select County",
                            choices = unique(bigfootDated$county), #instead of typing all options, we can use unique()
                            selected="Amador County"),
                selectInput("b", 
                            "Select Month", 
                            choices = c(1:12),
                            selected="1"),
                          
                selectInput("c", 
                            "Select day", 
                            choices = c(1:31), 
                            selected="1"),
                plotOutput(

                  "plot", width="500px", height="400px"),

), # close the first box    




box(title = "Plot of Penguins Data", width = 6, #adjust the column width and add title
  plotOutput("plot", width = "500px", height = "400px")
) # close the second box
) # close the row
) # close the dashboard body
) # close the ui

    
server <- function(input, output, session) {
  
output$plot <- renderPlot({
session$onSessionEnded(stopApp) # stop the app when we close it
  
  output$plot <- renderPlot({
  user_vector <- reactiveValues()
  #user_param<- input$a

  
value<-c(a,5,7)

value$a <- bigfootDated %>%
  count(county)%>%  
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter( county == "Amador County")%>%
  mutate(perc_CA_sightings = (n/total)*100)    

value$b <- bigfootDated %>%
  count(month)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter(month=="8")%>%
  mutate(perc_sightings_month = (n/total)*100)

value$c <- bigfootDated %>%
  count(day)%>%
  arrange(-n)%>% 
  mutate(total=sum(n))%>%
  filter(day== "7")%>%
  mutate(perc_sightings_day = (n/total)*100)




param<- c(input$a,input$b,input$c)
value<- c(value$a[4],value$b[4],value$c[4])

#trial %>% 
 # mutate(across(param, value, as.double))

print("PARAM")
print(param)
print("VALUE")
print(value)

param<-as.factor(param)
value<-as.numeric(value)


trial<- data.frame(param, value) 
class(param)
class(value)


      v<- trial%>%
          summarize(mean_val=mean(value))
print(trial) 

   
 
trial %>%
      ggplot(aes(x=param, y=value)) + 
      geom_col() +
         labs(x = "parameter", y = "percent of total sightings in california by parameter")+
     ggtitle( paste("Calculated chance of seeing bigfoot ", v, "%"))    

})
  
})

}

shinyApp(ui, server)
```

